---
title: "Data Visualization Project"
author: "Julian Lopez"
output:
  pdf_document: default
  html_document: default
---

# Data Visualization Project

## Objectives for this project

To complete this project you’ll need to do a few things within this file. As you go through the notebook, you will have further instruction on how to complete these objectives.

1. Go through this notebook, reading along.

2. Fill in empty or incomplete code chunks when prompted to do so.

3. Run each code chunk as you come across it by clicking the tiny green triangles at the right of each chunk. You should see the code chunks print out various output when you do this.

4. At the very top of this file, put your own name in the `author:` place. Currently it says `"DataTrail Team"`. Be sure to put your name in quotes.

5. In the `Conclusions` section, write up responses to each of these questions posed here.

6. When you are satisfied with what you've written and added to this document you'll need to save it. In the menu, go to `File` > `Save`. Now the `nb.html` output resulting file will have your new output saved to it.

7. Open up the resulting `airbnb_project.nb.html` file and click `View in Web Browser`. Does it look good to you? Did all the changes appear here as you expected.

8. Upload your `Rmd` and your `nb.html` files to your assignment folder (this is something that will be dependent on what your instructors have told you -- or if you are taking this on your own, just collect these projects in one spot, preferably a Google Drive)!

7. Pat yourself on the back for finishing this project!

## The goal of this analysis

"What variables relate to the cost of an Airbnb?”

## Set up

Let's load these packages for use. Add or subtract from this list as you see fit.

```{r setup, include=FALSE}
## load packages
library(janitor)
library(ggplot2)
library(dplyr)
library(ggthemes)
library(readr)
```


## Set up directories

Here we are going to make a `data` directory if it doesn't already exist.

```{r}
if (!dir.exists("data")) {
  dir.create("data")
}
```

## Get the data

The data we'll be using for this part of the project are about and can be downloaded from here: https://github.com/rfordatascience/tidytuesday/blob/master/data/2023/2023-01-24/readme.md

First, we'll read the data in from our `data/raw` directory. Use the `read_csv` function to do this. Call this new data frame `airbnb_df`.

```{r}
airbnb_df <- read_csv("/cloud/project/04_Plotting_Data/data/raw/Airbnb_Open_Data.csv")
```

Use this chunk to do some exploring of this dataset to get an idea of what kind of information is here.

```{r, solution = TRUE}
dim(airbnb_df)
ncol(airbnb_df)
str(airbnb_df) 
summary(airbnb_df)

## tells R to find the max amount in the price column and tells R to ignore NA in the data
max(airbnb_df$price, na.rm = T) 

## tells R to looks at the unique room types 
unique(airbnb_df$`room type`)

## tells you the number of unique room types 
length(unique(airbnb_df$`room type`))
```

### Clean the data

Use the `janitor::clean_names` function to clean up some of these column names.

```{r, solution = TRUE}
## lets take a look at the column names and see why they need to be cleaned
colnames(airbnb_df)

## overriding the bad column names
airbnb_df <- airbnb_df %>%
  janitor::clean_names()
```

### Fix `neighbourhood_group` variable

Take a look at the `neighbourhood_group` variable. Try to summarize how many different neighborhood groups there are using the `summary()` function.

```{r, solution = TRUE}
### summary('neighborhood_group)
### attempt one: summary(unique(airbnb_df$neighbourhood_group))
### using a factor can tell us the information we want see below
```

You will find that this variable is a character, so we will need to check it as a factor to get meaningful information about the groups here. Thus run the `summary` function on a factor version of the `neighborhood_group` data, but don't yet change the variable to be a factor.

```{r, solution = TRUE}
### change a categorical value to factor 
airbnb_df$neighbourhood_group <- as.factor(airbnb_df$neighbourhood_group)
### now use summary 
summary(airbnb_df$neighbourhood_group)

```

You'll notice there's some typos here `brookln` and `manhatan`. Fix these typos in this dataset by whatever method you find most intuitive. After you have fixed the typos, make sure that you coerce `neighbourhood_group` as a factor.

```{r, solution = TRUE}
### Helpful website to convert variables: https://sparkbyexamples.com/r-programming/replace-values-in-r/

airbnb_df$neighbourhood_group[airbnb_df$neighbourhood_group == 'brookln'] <- 'Brooklyn'
airbnb_df$neighbourhood_group[airbnb_df$neighbourhood_group == 'manhatan'] <- 'Manhattan'

```

Re-run `summary()` on the `neighbourhood_group` to see if your code successfully fixed these misspellings.

```{r, solution = TRUE}
summary(airbnb_df$neighbourhood_group)
```

### Fix numeric data

The `service_fee` and `price` columns are really numeric data but are being treated as characters because there are `$` and `,` included.

We'll need to get rid of these. **Hint** use this kind of code:
`stringr::str_remove(service_fee, "\\$|,")` within a mutate to get rid of the dollar signs. After you get rid of the dollar signs from both `service_fee` and `price`, coerce both of these variables to be numeric.

```{r, solution = TRUE}
?stringr::str_remove()
airbnb_df <- airbnb_df%>%
  mutate(service_fee = stringr::str_remove(service_fee, "\\$|,"))

airbnb_df <- airbnb_df%>%
  mutate(price = stringr::str_remove(service_fee, "\\$|,"))
```

Use summary to see if `price` and `service_fee` are now appropriately numeric data.

```{r, solution = TRUE}
airbnb_df$service_fee <- as.numeric(airbnb_df$service_fee)
```

```{r, solution = TRUE}
airbnb_df$price <- as.numeric(airbnb_df$price)
```

## Plot the data

Let's make a series of plots that explore how the price of an airbnb may be related to the other variables in this set.

To refresh your memory, a basic ggplot code looks like this:

```
DATA_FRAME %>%
  ggplot(aes(VARIABLE, VARIABLE, color = OPTIONAL_VARIABLE)) +
  geom_PLOT_TYPE()
```

We recommend keeping the ggplot2 cheatsheet handy while you are going through this project:
https://www.rstudio.com/resources/cheatsheets/

### Scatter plot

Now make a scatter plot that explores if `price` is related to one of the other numeric variables in the dataset.

```{r, solution = TRUE}
  ggplot(data = airbnb_df) +
  geom_point(mapping = aes(x = price, y = service_fee))
### Relationship observed on graph: As the price increases so does the service fee
```

### Violin plot

Make a violin plot for the price distribution of different neighborhood groups.

```{r, solution = TRUE}
ggplot(data = airbnb_df) +
  geom_violin(mapping = aes(y = price, x = neighbourhood_group))
```

### Boxplot

Use a box plot for price and neighborhood groups relationship.

```{r}
ggplot(data = airbnb_df) +
  geom_boxplot(aes(y = price, x = neighbourhood_group), fill = "purple")+
  scale_y_continuous(breaks = seq(0,250, by = 10))
```


### Barplot

Now make a plot that explores if there's a particular `room_type` that has a higher price than others. You may choose to summarize the `price` by the `room_type`.  Hint: you may need to `group_by` a variable.    

```{r, solution = TRUE}
airbnb_df %>%
  group_by(room_type) %>%
  summarize(pricenew = mean(price, na.rm = T)) %>%
  ggplot(aes(x=room_type, y = pricenew)) +
  geom_bar(fill = "green", stat = "identity")
```

### Your choice plot!

Make one more plot that is completely your choice about what variables you'd like to see the relationships of.

See http://r-statistics.co/Top50-Ggplot2-Visualizations-MasterList-R-Code.html for inspiration and ideas.

Looking to see the relationship between the two types of host identities, verified and unverified, and it effect on the hosts rating score. 

```{r, solution = TRUE}
airbnb_df %>%
  group_by(host_identity_verified) %>%
  summarize(rating_mean = mean(review_rate_number, na.rm = T)) %>%
  ggplot(aes(x = host_identity_verified, y = rating_mean))+
  geom_bar(fill = "red", stat = "identity")
```

### Conclusion

Write up your thoughts about this data science project here and answer the following questions:

- What did we find out about our questions?
Of the variables we explored, there is a positive linear relationship between the service fee and price of an airbnb. 
As the service fee increased so to did the price of the airbnb. 

- How did we explore our questions?
We explored our questions by incorporating plots displaying the relationship between two different variables. 

- What did our explorations show us?
Our explorations yielded few significant relatiosnhips with the exception of the relaioship between the service fee and price of an airbnb

- What follow up data science questions arise for you regarding this airbnb dataset now that we've explored it some?
I would like to see the relationship between the construction date and the rating of the airbnb. 
I would like to see if the newer the airbnb in terms of construction date the higher airbnb rating. 

## Print out session info

Session info is a good thing to print out at the end of your notebooks so that you (and other folks) referencing your notebooks know what software versions and libraries you used to run the notebook.

```{r}
sessionInfo()
```
